---
title: "Final"
output:
  pdf_document: default
  html_document: default
date: "2023-11-28"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


Dataset:

@misc{h-and-m-personalized-fashion-recommendations,
    author = {Carlos García Ling, ElizabethHMGroup, FridaRim, inversion, Jaime Ferrando, Maggie, neuraloverflow, xlsrln},
    title = {H&M Personalized Fashion Recommendations},
    publisher = {Kaggle},
    year = {2022},
    url = {https://kaggle.com/competitions/h-and-m-personalized-fashion-recommendations}
}


```{r}
require(plyr)
require(dplyr)
require(tidyverse)
require(haven)
library(readr)
```

Summary of data

Articles
  articles of clothing
  article id and decriptions
  
Customers
  customer id, age and club membership
  
Transaction
  customer id, article id, price, and amount of sales


## Loading Data


```{r}
#articles <- read_csv("/Users/vanessacrowe/Desktop/Econometrics R Projects/articles.csv")

#Subset with deleted columns in Articles of clothing (deleting produce code/name and perceived values for customers) and omitting na values

#articlesset <- articles[, -c(2, 3, 11, 12, 13, 14, 15, 16, 17, 18, 19, 20, 21, 22, 23, 24)]
#articlesset <- na.omit(articlesset)

#print(articlesset)
#View(articlesset)

```


```{r}
#customers <- read_csv("/Users/vanessacrowe/Desktop/Econometrics R Projects/customers.csv")

#Delete irrelevant columns from customers and omitting na values

#customersset <- customers[, -c(2, 3, 5, 7)]
#customersset <- na.omit(customersset)

#print(customersset)
#View(customersset)

```

```{r}
# deleted after saved new smaller file
#transactions_train <- read_csv("/Users/vanessacrowe/Desktop/Econometrics R Projects/transactions_train.csv")

#Cutting to 12.5% of the training data to only 3.7 million observations (from 31 million) and omitting na just in case

#set.seed(123)

#sampletransations <- transactions_train %>%
  #sample_n(size = 0.125 * nrow(.))

#sampletransations <- na.omit(sampletransations)
#View(sampletransations)

```

Save as csv
```{r}
#write.csv(sampletransations, file = "/Users/vanessacrowe/Desktop/Econometrics R Projects/sampletransactions.csv")
```

```{r}
#Transactions <- read_csv("/Users/vanessacrowe/Desktop/Econometrics R Projects/sampletransactions.csv")
#View(Transactions)
```
This should make it easier to identify consumer preferences 
Merge data to include even unmatched rows first by customer ids since the two data sets share that column

```{r}

#customerstransactions <- merge(customersset, Transactions, by = "customer_id", all = TRUE)

#Merge everything else by article id
#HandMTransactions <- merge(customerstransactions, articlesset, by = "article_id", all = TRUE)

```

get rid of na  values again and the added column counting the values
```{r}

#HandMTransactions <- na.omit(HandMTransactions)
#HandMTransactions <- HandMTransactions[, -c(5)]
#View(HandMTransactions)

```

```{r}
#write.csv(HandMTransactions, file = "/Users/vanessacrowe/Desktop/Econometrics R Projects/HandMTransactions.csv")

```


```{r}
#head(HandMTransactions)
```

Maybe most popular season is Spring and Summer
April - September vs October - March

First date: September 20, 2018 - September 22, 2020

Separate by dates with similar clothing options

Fall/Winter Months
September 2018 - March 2019
October 2019 - March 2020

Spring/Summer Months
April 2019 - September 2019
April 2020 - September 2020

There are a lot more small peaks in this graph for Spring and Summer while the FW season had a lot more extreme peaks in sales.

Subset of only Fall and Winter Months

```{r}
#start_date1 <- as.Date("2018-09-20")
#end_date1 <- as.Date("2019-03-31")

#start_date2 <- as.Date("2019-09-23")
#end_date2 <- as.Date("2020-03-31")

#HMTransactionsFW <- subset(
#  HandMTransactions,
#  (t_dat >= start_date1 & t_dat <= end_date1) |
#  (t_dat >= start_date2 & t_dat <= end_date2)
#)

#print(HMTransactionsFW)

# 1,873,867 transactions

```

Find out how much sales they are making each date
Count the items (non specific garments to how much is sold on certain days)

```{r}
#summary <- HMTransactionsFW %>%
#  group_by(t_dat) %>%
#  count()
#head(summary)
```

Most transactions for FW months was 2019-09-28 with 24,587

```{r}
#max_rowFW <- summary[which.max(summary$n), ]
#max_rowFW
```


```{r}
library(dplyr)
library(ggplot2)
library(plotly)

#PlotFW <- ggplot(summary, aes(x = t_dat, y = n, group = 1)) +
 # geom_line(color = "blue") +
#  labs(title = "Sales by Days Fall/Winter",
    #  x = "Date",
     #  y = "Sales") +
 # theme_minimal()
#scale_x_date(date_breaks = "2 months", date_labels = "%Y-%m")

#PlotFW <- ggplotly(PlotFW)

#PlotFW
```
Peaks:
2018-09-29 - 17,364, no holiday but close to the beginning of fall
2018-11-23 - 17,677, Black Friday
2019-11-29 - 20,122, Black Friday

2019-02-23 - 9,409, midwinter recess
2019-09-28 - 24,587 no holiday but close to the beginning of fall
2019-11-03 - 10,599 
2019-11-28 - 13,205 -Thanksgiving
2019-12-18 - 8,188 
2020-03-04 - 7,049

Lowest:
2018-09-22 - 2,225
2018-12-25 - 1,843 - Christmas
2019-01-01 - 2,256 - New Years

2019-12-25 - 1,653 - Christmas
2020-01-01 - 1,544, New Years


FW: add column price based on same clothing article and group by each month, now if they want to determine how much they made each month from each article of clothing, they can used the 'articles_purchased' column to multiply with the 'total_price'

```{r}

#SalesbyPrice <- HMTransactionsFW %>%
 #   group_by(article_id, month = format(t_dat, "%Y-%m")) %>%
 # summarise(total_price = sum(price), articles_purchased = n())

#print(SalesbyPrice)
```
Highest and lowest purchases months
Least purchases in February and most purchases in November

```{r}
#SalesbyPriceFW <- SalesbyPrice$total_price * SalesbyPrice$articles_purchased
#lowest_row <- SalesbyPrice[which.min(SalesbyPriceFW), ]
#highest_row <- SalesbyPrice[which.max(SalesbyPriceFW), ]
#lowest_row
#highest_row
```


Subset of only Summer and Spring Months

```{r}

#start_date3 <- as.Date("2019-04-01")
#end_date3 <- as.Date("2019-09-22")

#start_date4 <- as.Date("2020-04-01")
#end_date4 <- as.Date("2020-09-22")

#HMTransactionsSS <- subset(
#  HandMTransactions,
#  (t_dat >= start_date3 & t_dat <= end_date3) |
 # (t_dat >= start_date4 & t_dat <= end_date4)
#)

#print(HMTransactionsSS)

# 2,062,015 transactions

```

```{r}
#summary1 <- HMTransactionsSS %>%
#  group_by(t_dat) %>%
#  count()
# head(summary1)
```

Most transactions during the SS months - 2019-04-11 of 20,386
```{r}
#max_rowSS <- summary1[which.max(summary1$n), ]
#max_rowSS
```


```{r}

# PlotSS <- ggplot(summary1, aes(x = t_dat, y = n, group = 1)) +
#   geom_line(color = "orange") +
#   labs(title = "Sales by Days Spring/Summer",
 #       x = "Date",
#        y = "Sales") +
#   theme_minimal()
# scale_x_date(date_breaks = "2 months", date_labels = "%Y-%m")

# PlotSS <- ggplotly(PlotSS)

## PlotSS
 
```
Peaks:
2019-04-06 - 14,534
2019-04-24 - 9,912
2019-05-18 - 12,050, leading up to memorial day
2019-06-19 - 12,769
2019-06-25 - 11,388
2019-06-26 - 11,837
2019-07-24 - 15,452

Consistently a lot of sales in mid June could be related to summer shopping and end of school celebration
2020-04-04 - 10,063, spring break
2020-04-11 - 20,386, spring break, day before easter
2020-06-17 - 13,875
2020-06-18 - 10,562
2020-06-19 - 9,836
2020-06-24 - 12,131
2020-06-25 - 10,926
2020-06-26 - 9,635
2020-08-12 - 8,691, back to school

Lowest:
2019-04-27 - 3,559
2019-06-09 - 3,676
2019-09-21 - 2,524
2020-05-03 - 3,122
2020-08-16 - 2,845

SS: add column price based on same clothing article and group by each month, now if they want to determine how much they made each month from each article of clothing, they can used the 'articles_purchased' column to multiply with the 'total_price'

```{r}

#SalesbyPrice1 <- HMTransactionsSS %>%
 #    group_by(article_id, month = format(t_dat, "%Y-%m")) %>%
#   summarise(total_price = sum(price), articles_purchased = n())

# print(SalesbyPrice1)
 
```
Most and least purchases were made in June 2019 but they were different articles of clothing
```{r}
#SalesbyPriceSS <- SalesbyPrice1$total_price * SalesbyPrice1$articles_purchased
#lowest_row1 <- SalesbyPrice1[which.min(SalesbyPriceSS), ]
#highest_row1 <- SalesbyPrice1[which.max(SalesbyPriceSS), ]
#lowest_row1
#highest_row1

```

Dummy variables: created to assign a numerical value to a categorical variable between 0 and 1. Each represents one category of an explanatory or response variable (what you manipulate or observe changes in) coded with 1 if it is and 0 if it is not. For example, female = 1 is true and female = 0 is false as indicators. 

What days of the week gets the most purchases?


```{r}
#HandMTransactions <- HandMTransactions %>%
#  mutate(day_of_week = weekdays(t_dat),
   #      year = format(t_dat, "%Y"))
#head(HandMTransactions)

```


```{r}
#HandMTransactions <- HandMTransactions %>%
#  mutate(day_of_week = as.factor(day_of_week),
  #       year = as.factor(year))
#head(HandMTransactions)

```


The most popular day of purchase was Saturday with 625,677 purchases

```{r}
#day_counts <- HandMTransactions %>%
#  group_by(day_of_week) %>%
#  summarise(count = n())
#day_counts

```

```{r}
#HMTransactionsFW <- HMTransactionsFW %>%
#  mutate(day_of_week = weekdays(t_dat),
 #        year = format(t_dat, "%Y"))
#head(HMTransactionsFW)

```


```{r}
#HMTransactionsFW <- HMTransactionsFW %>%
 # mutate(day_of_week = as.factor(day_of_week),
  #       year = as.factor(year))
#head(HMTransactionsFW)

```


The most popular day of purchase for Fall and Winter was Saturday with 313,447 purchases

```{r}
#day_countsFW <- HMTransactionsFW %>%
 # group_by(day_of_week) %>%
 # summarise(count = n())
#day_countsFW

```

```{r}
#HMTransactionsSS <- HMTransactionsSS %>%
  #mutate(day_of_week = weekdays(t_dat),
   #      year = format(t_dat, "%Y"))
#head(HMTransactionsSS)

```


```{r}
#HMTransactionsSS <- HMTransactionsSS %>%
#  mutate(day_of_week = as.factor(day_of_week),
  #       year = as.factor(year))
#head(HMTransactionsSS)

```


The most popular day of purchase for Spring and Summer was Saturday with 336,137 purchases

```{r}
#day_countsSS <- HMTransactionsSS %>%
 # group_by(day_of_week) %>%
 # summarise(count = n())
#day_countsSS
```

Convert to numeric so now all values called 'ACTIVE' is 1 and all others(PRE-CREATE and LEFT CLUB) is 0 for false 

```{r}
#HandMTransactions$club_member_status <- as.numeric(HandMTransactions$club_member_status == "ACTIVE")
```

#name for active memnbers in the club
```{r}
#HandMACTIVE <- as.numeric(HandMTransactions$club_member_status == "ACTIVE")
#table(HandMTransactions$club_member_status)

#3857870 active members and 78012 inactive members (leftclub or pre-created account)
```

Create a subset with the specific factors from the original data set - target age range of H&M is 0-40 and most popular buyers range from 15-30 but it is a little hard for babies to make purchases
```{r}
#use_data <- (HandMTransactions$age >= 15) & (HandMTransactions$age <= 40)
#HMSubset <- subset(HandMTransactions,use_data)
#HMSubset

# from 3,935,882 to 2,556,372
```

Subset to combine all the duplicated customer_ids and create a column of the quantity adding up the prices to use to compare - this will also include the other columns. How much each customer is buying? organize by unique customer_id's (if article is different, that is counted)

```{r}

#HMSubset <- HMSubset %>%
   #group_by(customer_id) %>%
 #  mutate(customer_count = n(), total_price1 = sum(price))
```

```{r}
#head(HMSubset)
```
Count customer_count to see what customer purchases the most - customer count and age
```{r}
#ranges <- cut(HMSubset$age, breaks = c(15, 20, 25, 30, 35, 40, labels = c("15-20", "21-25", "26-30", "31-3#5", "36+"), include.lowest = TRUE))
#HMSubset$ranges <- ranges
##head(HMSubset)
```

```{r}
#purchase_by_age_range <- HMSubset %>%
  #group_by(ranges) %>%
 # summarise(customer_count = sum(customer_count))
```

The age range that purchased the most was [25,30] with 11,334,707	items in total.

```{r}
#purchase_by_age_range
```
```{r}
#ggplot(HMSubset, aes(x = age, y = customer_count)) +
#  geom_point() +
#  labs(title = "Scatter Plot of Customer Count vs. Age")

```


```{r}
#summary_stats <- summary(HMSubset[, c("customer_count", "age", "product_type_no", "club_member_status")])
#print(summary_stats)
```


How do all of those variables affect amount customer buys?
Relationship between the dependent variable customer_count and independent variables age, product number, and club member status
Predicted values of customer count when all independent variables are zero (intercept)

Cannot regress price on quantity, in this case it is 'customer_count'

```{r}
#ols_model <- lm(customer_count ~ age + product_type_no + club_member_status,  data =  HMSubset)
#summary(ols_model)
```
Based on the values of age, price, product number, club status, withtin the age range of 15-40, what is the main determiner of how much they are willing to buy?

For each one-unit increase in 'age,' the model predicts an increase of approximately 0.3621 in the 'customer_count,' holding other variables constant. As customer gets older, the more they are purchasing. 
The effects of product_type_no on customer_counts is not that important. This means that as the products go from small accessories or outer summer wear like a hair clip(72), belts (67), top and lower garmets (253) to items like long socks (302), stockings (304), the customer is less likely to purchase 
For each one-unit increase in 'club_member_status' (assuming it's binary), the model predicts an increase of approximately 8.815 in 'customer_count.



```{r}
# subset in order to plot...
#calculates the number of observations in the 'customer_count' column in the HMsubset column
#NNobs <- length(HMSubset$customer_count)
#set.seed(12345) # just so you can replicate and get same "random" choices
#graph_obs <- (runif(NNobs) < 0.1) # so something like just 1/10 as many obs
#dat_graph <-subset(HMSubset,graph_obs)  

#plot(customer_count ~ jitter(age, factor = 2), pch = 16, col = rgb(0.5, 0.5, 0.5, alpha = 0.2), data = dat_graph)
#plot(customer_count ~ jitter(age, factor = 2), pch = 16, col = rgb(0.5, 0.5, 0.5, alpha = 0.2), ylim = c(0,50000), data = dat_graph)
# discus what you see in this plot

# change this line to fit your choices about explanatory variables
#to_be_predicted2 <- data.frame(age = 15:40, price = 0.016932203, product_type_no = 253, club_member_status = 1)
#used to predict this specific values for the predictor variable (customer_count)
#to_be_predicted2$yhat <- predict(ols_model, newdata = to_be_predicted2)

#lines(yhat ~ age, data = to_be_predicted2)
#scatter plot shows the relationship between age and customer count for a subset of data

```
```{r}
require(tidyverse)
require(class)
require(caret)
```

```{r}
# Assuming HMSubset is your original data frame
# Set seed for reproducibility
#set.seed(12345)

# Number of observations
##NN_obs3 <- nrow(HMSubset)

# Create a binary indicator for training data (80% for training, 20% for testing)
#select1 <- (runif(NN_obs3) < 0.8)

# Split the data into training and testing sets
#train_data <- HMSubset[select1, ]
#test_data <- HMSubset[!select1, ]
#cl_data <- train_data$customer_count
#true_data <- test_data$customer_count

# Summary of training and testing data
#summary(cl_data)
#summary(train_data)

```
Lowest price: 0.0000508
Highest price: 0.5915254
There is an overwhelming majority of [20,25] and [25,30] year olds buying from h&m almost double the amounts from other age ranges


What is the most popular price? or what is the most popular item? 
Different prices are different items so by finding the price counts, you can find what items are the most popular

```{r}
#price_counts <- HMSubset %>%
#  group_by(price) %>%
 # summarise(count = n())
```


```{r}
#price_counts <- price_counts %>%
 # arrange(desc(count))
#head(price_counts)
#tail(price_counts)
```
Quite a bit of outliers which are the most common clothing pieces customers buy or the least common. Showing the customers prioritize certain items more than others and it is not a widespread and the concentrated areas is more so because the prices of the items are so close to each other. Only a few items are even above 10k sales

```{r}
#scatter_plot <- plot_ly(price_counts, x = ~price, y = ~count, type = 'scatter', mode = 'markers', marker = list(color = 'red'))
#scatter_plot
```


```{r}
#top_10_prices <- head(price_counts, 10)
#print(top_10_prices)
```
Surprisingly, the prices with the most sales are the not the lowest prices but they are tied to items that are typically used as 'staples' or 'basics' like the most popular price had 272,937 sales which is a white microfiber t shirt bra

Could be because the prices are close to each other.


Now predicting the price/product bought for a specific consumer
```{r}
#ols_model2 <- lm(price ~ age + product_type_no + club_member_status,  data =  HMSubset)
#summary(ols_model2)
```
For each one-unit increase in 'age,' the model predicts an increase of approximately 0.0001479 in 'price,' holding other variables constant.
For each one-unit increase in 'product_type_no,' the model predicts an increase of approximately 0.00002191 in 'price.' his means that as the products go from small accessories or outer summer wear like a hair clip(72), belts (67), top and lower garmets (253) to items like long socks (302), stockings (304), the customer is less likely to purchase
For each one-unit increase in 'club_member_status' (assuming it's binary), the model predicts a decrease of approximately 0.001717 in 'price.
The model overall explains a relatively small proportion of the variability in the price


What colors are the most popular and by what seasons and what styles are the most popular?


Fall and Winter
```{r}

# Most popular 'colour_group_name'
#popular_colour_groupFW <- table(HMTransactionsFW$colour_group_name)
#popular_colour_groupFW <- popular_colour_groupFW[order(-popular_colour_groupFW)]
#print(head(popular_colour_groupFW, n = 10))

# Most popular 'detail_desc'
#popular_detail_descFW <- table(HMTransactionsFW$detail_desc)
#popular_detail_descFW <- popular_detail_descFW[order(-popular_detail_descFW)]
#print(head(popular_detail_descFW, n = 10))

# Most popular 'product_type_name'
#popular_product_typeFW <- table(HMTransactionsFW$product_type_name)
#popular_product_typeFW <- popular_product_typeFW[order(-popular_product_typeFW)]
#print(head(popular_product_typeFW, n = 10))

```

```{r}
# Least popular 'colour_group_name' in HMTransactionsFW
#least_popular_colour_groupFW <- table(HMTransactionsFW$colour_group_name)
#least_popular_colour_groupFW <- least_popular_colour_groupFW[order(least_popular_colour_groupFW)]
#print(head(least_popular_colour_groupFW, n = 10))  # Using head to show the least popular

# Least popular 'detail_desc' in HMTransactionsFW
#least_popular_detail_descFW <- table(HMTransactionsFW$detail_desc)
#least_popular_detail_descFW <- least_popular_detail_descFW[order(least_popular_detail_descFW)]
#print(head(least_popular_detail_descFW, n = 10))  # Using head to show the least popular

# Least popular 'product_type_name' in HMTransactionsFW
#least_popular_product_typeFW <- table(HMTransactionsFW$product_type_name)
#least_popular_product_typeFW <- least_popular_product_typeFW[order(least_popular_product_typeFW)]
#print(head(least_popular_product_typeFW, n = 10))  # Using head to show the least popular

```


Spring and Summer

```{r}
# Assuming your data is stored in a data frame named 'your_data'

# Most popular 'colour_group_name'
#popular_colour_groupSS <- table(HMTransactionsSS$colour_group_name)
#popular_colour_groupSS <- popular_colour_groupSS[order(-popular_colour_groupSS)]
#print(head(popular_colour_groupSS, n = 10))

# Most popular 'detail_desc'
#popular_detail_descSS <- table(HMTransactionsSS$detail_desc)
#popular_detail_descSS <- popular_detail_descSS[order(-popular_detail_descSS)]
#print(head(popular_detail_descSS, n = 10))

# Most popular 'product_type_name'
#popular_product_typeSS <- table(HMTransactionsSS$product_type_name)
#popular_product_typeSS <- popular_product_typeSS[order(-popular_product_typeSS)]
#print(head(popular_product_typeSS, n = 10))

```

```{r}
# Least popular 'colour_group_name' in HMTransactionsFW
#least_popular_colour_groupSS <- table(HMTransactionsSS$colour_group_name)
#least_popular_colour_groupSS <- least_popular_colour_groupSS[order(least_popular_colour_groupSS)]
#print(head(least_popular_colour_groupSS, n = 10))  

# Least popular 'detail_desc' in HMTransactionsFW
#least_popular_detail_descSS <- table(HMTransactionsSS$detail_desc)
#least_popular_detail_descFW <- least_popular_detail_descSS[order(least_popular_detail_descSS)]
#print(head(least_popular_detail_descSS, n = 10)) 

# Least popular 'product_type_name' in HMTransactionsFW
#least_popular_product_typeSS <- table(HMTransactionsSS$product_type_name)
#least_popular_product_typeSS <- least_popular_product_typeSS[order(least_popular_product_typeSS)]
#print(head(least_popular_product_typeSS, n = 10))  

```






